Synergistic retrievals of ice in high clouds from elastic backscatter lidar, Ku-band radar and submillimeter wave radiometer observations

### 2023-05-16 RST/responseReviewers.tex

they are consistent with an observed vertical distribution of reflectivity at W-band. The vertical distributions of radar reflectivity
derived from satellite or ground-based observations are generally difficult to reproduce by model simulations.
We added statements
to this effect in the revised manuscript (lines 67-69 in the Introduction).\\

@@

For consistency and simplicity in the revised manuscript we derive IWC-Z relationships 
for each CloudSat radar bin, with the bins indexed relative to the freezing level bin.  The relative height is defined as the distance
between the height of the associated bin relative to the freezing level.  Results are rather similar, although the overall performance is
degraded to some degree due to the inclusion of the $N_w$ perturbation component.\\

@@

data in Fig. 2? This needs to be clarified.}\\
\newline
Yes, we fit a line to determine the variation of $ln(a)$ as a function of height and then divide the slope by $(1-b)$ to get the variation of
$ln(N_w)$ as a function of height. We added explanations to this effect in the revised manuscript (lines 114-124).\\

@@

we randomly perturb $N_w$ by multiplying its parameterized value with log-normally distributed
random variable with zero mean a standard deviation of 0.5.  As explained above, we use a Gaussian smoothing filter to impose a
vertical auto-correlation in
$N_w$ similar to that observed in field experiments. This is explained in lines 155-165 of the revised manuscript.\\

@@

of the ($Z/N_w$,$IWC/N_w$) distribution to almost one-to-one relationships.  As explained in the manuscript, we use the self-similar Rayleigh-Gans approximation
to model the backscattering property of ice particles but get a normalization behavior similar to that anticipated by Testud et al. (2001),
shown in \cite{delanoe2014} and exhbited in the \cite{chase2021} database.
We added statements to this effect in the paragraphs describing Equation (1).\\

@@

estimation process more robust. However, for consistency and simplicity in the revised manuscript we derive IWC-Z relationships 
for each CloudSat radar bins, with the bins indexed relative to the freezing level bin.  The relative height is defined as the distance
between the height of the associated bin relative to the freezing level.  Results are rather similar, although the overall performance is
degraded to some degree due to the inclusion of the $N_w$ perturbation.\\

### 2023-05-16 RST/RSTManuscript.tex
Review May 16

To achieve its objective, AOS will rely on a combination of active and passive observations onboard multiple spacecrafts

@@

complimentary for synergistic characterization of ice clouds. 

@@

good first step towards the development of an operational synergistic retrieval of ice clouds from the AOS inclined observations.

@@

\subsection{Assumptions and forward models}
To quantify the number of ice particles in an elementary atmospheric volume as a function of their size, we use
normalized gamma functions \citep{bringi2003}.  The benefit of normalized gamma functions is that they 
encapsulate the variability of Ice Water Content (IWC) - reflectivity relationship into a single parameter,
i.e. the 
normalized Particle Size Distribution (PSD) intercept \citep{testud2001,bringi2003}. The normalized
PSD intercept is defined as $N_w=\frac {4^4} {\pi \rho_w} \frac {IWC} {D_m^4}$, where $IWC$ is the ice water 
content associated with the PSD, and $D_m$ is the mass weighted mean diameter.  \cite{testud2001} showed 
that the variability in IWC reflectivity (Z) relationships may be fully explained by variability in $N_w$, and 
that a formula of the type
\begin{equation}
IWC=N_w^{1-b}a_{norm} Z^b \label{eq:1}
\end{equation}
perfectly explains the relationships between IWC and Z calculated from observed PSDs. 
Although the initial work of \cite{testud2001} was limited to rain and only hinted about the applicability
of the normalization concept to ice observations, subsequent studies showed that parameterizations
of type (\ref{eq:1}) extended to ice \citep{ferreira2001,delanoe2014}.
However, equation (\ref{eq:1}) is not 
sufficient to derive accurate, unbiased estimates of ice water contents, because $N_w$ varies considerably in 
time and space. Nevertheless, multiple studies showed that it is beneficial to parameterize $N_w$ as a function 
of various variables, such as temperature \citep{hogan2006retrieval, delanoe2008variational,deng2010tropical}, 
rather than using $N_w$ independent reflectivity ice relations.  In this study, we parameterize
$N_w$ as a function of temperature based on the CloudSat 2C-ICE product \citep{deng2010tropical,deng2013evaluation}.  
Specifically, we cluster, based on similarity (quantified through the evaluation of the Euclidian distance between profiles), 
a large set 2C-ICE profiles into 18 classes using a k-Means 
procedure. The mean IWC profiles associated with the 18 classes are shown in continuous lines in Figure 1.  
Our estimates, derived using PSD assumptions and electromagnetic scattering calculations that enable 
accurate and physically consistent simulations of radar observations at Ku-band and radiometer observations of 
submillimeter-wave frequencies are also shown in Figure 1. These estimates are based on the self-similarity 
Rayleigh-Gans approximation (SSRGA) of \cite{hogan2017ssrga}. Details regarding the
estimation process are provided in the following paragraphs.  As apparent in Figure 1, the CS and SSRGA estimates are in good 
agreement.  Some discrepancies due to differences between the SSRGA $N_w$ parameterization and the CS 2C-ICE 
"a priori assumptions" are also apparent, but they are not deemed critical in this study, whose objective is 
the investigation of synergistic lidar, Ku-band radar and submillimeter-wave radiometer retrievals, because the 
outcome is not likely to be sensitive to such details.
\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig01.png}\\
    \caption{Mean CS IWC profiles for 18 classes derived using the k-Means clustering algorithm. For a 
    compressed but intelligible representation, three classes are shown in different colors in each panel.
    The associated 
    mean profiles derived from CS reflectivity observations using the SSRGA scattering calculations 
    and $N_w$ parameterization developed in this study are shown using symbol *. The vertical coordinate 
    is defined relative to the freezing level.}\label{f1}
\end{figure}

A scatter-plot analysis (not shown) of relationships between the 2C-ICE IWC and the associated reflectivity suggests that
$a$ is parameterized as a function of temperature (or equivalently, as a function of height relative to the freezing level) in 
the default 2C-ICE retrievals. The default $a$ (the value the provides the IWC estimate prior to the ingestion of the lidar
observations) may be simply estimated by regressing the 2C-ICE IWC on the associated CloudSat reflectivity as a function of 
height relative the freezing level. Specifically, a regression of the type $ln(IWC)=ln(a)+b*ln(Z)$ is performed for 
each bin relative to the zero degree bin is performed, and variation of $ln(a)$ is parameterized as a function of the
relative height.  The results are shown in Figure 2. As apparent in the figure, and as expected, $a$ exhibits a strong variation 
with the relative height.  Coefficient $b_i$ exhibits a height dependency as well (not shown), but the range of variation is 
significantly smaller, almost zero relative to the mean value of $b$. Given that any deviation of the multiplicative coefficient in 
an IWC-Z relation from an average is equivalent to a deviation of the associated $N_w$ from its mean value 
(Testud et al. 2001), the variation of $a$ as a function of relative-height may be converted into a $N_w$ 
as a function of relative-height relationship.  We, therefore, use the data in Fig. 2 to parameterize $N_w$ as a 
function of the relative height.  Specifically, given that $(1-b)ln(N_w)+ln(a_{norm})=ln(a)$ and $ln(a)=ln(a_0)+s*H$, we 
can express $N_w$ as a function of the relative height, slope $s$, and parameters $a_0$ and $a_{norm}$.

@@

For the determination of reference $a$ and $b$ values to be used with Equation (\ref{eq:1}), we assume that 
PSDs are normalized gamma  distributions with $N_w=0.08cm^{-4}$ and $\mu=2$ and calculate 
\begin{equation}
 Z=\frac {\lambda ^4} {\pi ^5 |K_w|^2} \int_0^{\infty} N(D,D_m) \sigma _b(D) dD
\end{equation}
where $\lambda$ is the radar frequency, $|K_w|$ is the dielectric factor of water, $N(D,D_m) dD$ is the number 
of ice particles of diameter with D and D+dD per unit volume, $D_m$ is the mass weighted mean diameter of the 
distribution, and $\sigma _b(D)$ is the backscattering cross-section of ice particle of diameter $D$. The mass 
weighted mean diameter is equidistantly sampled to span the entire range of IWC values in the CS 2C-ICE dataset.
The assumed mass-size relation is that of \cite{brown1995improved} because it works well with the 
SSRGA scattering calculations \citep{heymsfield2022impacts}.  The open source software scatter-1.1 of \citep{scatter-1.1}
is used to provide the actual scattering properties.   
To improve the representation of microphysical variability in the study, we do not assume the values of $N_w$ given
by the 2C-ICE based parametrization described above as the true values. Instead, we perturb them by multiplication 
with a log-normally distributed random variable with 0.0 mean a standard deviation of 0.5.  The 
perturbations are vertically correlated. Specifically, normal random variables with zero mean and a standard deviation
of 1.0 are generated and then passed through a Gaussian smoothing filter \citep{nixon2019} with a size of two radar bins. The perturbation
vector is then rescaled to have a standard deviation of 0.5.  The perturbations are then exponentiated and multiplied
with the 2C-ICE based $N_w$ values to produce the final $N_w$ values.  The filter size and noise magnitude are chosen
to roughly mimic the observed variability in $N_w$ derived from in situ observations of PSDs, such as those described
in \cite{heymsfield2022impacts}. In addition, we are adding a random noise of
0.0 mean and 0.5 dB standard deviation to the calculated reflectivity values, which is consistent with the expected
performance of a space-borne Ku-band radar system \citep{takahashi2008}.  Nevertheless, it should be noted that the
uncertainties in the reflectivity calculations are likely to be even larger, given that the SSRGA theory although 
quite accurate cannot possibly capture the entire range of uncertainties in the scattering properties of ice particles.
Specifically, the SSRGA calculations were carried out assuming ice particles consisting of aggregates of 
bullet rosettes, columnar crystals, and plates.

@@

The radiometer observations are calculated using a one-dimensional efficient, but accurate, radiative transfer 
solver based on Eddington's approximation \citep{kummerow1993edd}. The Eddington's approximation has been found to work 
well in cloud and precipitation retrieval application despite its simplicity relative to more general (but also 
computationally intensive) approaches such as Monte Carlo radiative transfer solvers \citep{liu1996three} 
It should be noted tough that the phase functions of ice particles tend to be highly asymmetric at 
sub-millimeter wave frequencies. For radiative transfer solutions based on the Eddington's approximation to be 
accurate, it is necessary that the delta-scaling approach \citep{joseph1976delta} be employed. The delta-scaling 
approach transforms the initial radiative transfer equation into an equivalent one characterized by a less 
asymmetric scattering function and more extinction, which makes the solution Eddington approximation more stable 
and accurate. The absorption due to water vapor and other gases is quantified using the Rosenkranz model 
\citep{rosenkranz1998}.  The water vapor, temperature, and pressure distributions are derived based on a WRF 
simulation of summer convection over the United States.  Specifically, the water vapor, temperature, and 
pressure profiles associated with times and areas where the model produces anvils are selected and clustered 
into 40 classes using the k-Means approach.  The mean extinction profiles at the radiometer frequencies are 
calculated for every class 
and used in process of calculating the brightness temperatures from the estimated ice profiles using a simple 
Monte Carlo procedure.  That is, given a retrieved ice profile and its scattering property, an anvil class 
and its associated absorption, temperature, and pressure profiles are randomly selected and attached to the 
ice scattering properties. To make the procedure physically meaningful, temperature rather than height is 
used in the ice scattering-gas absorption collocation process.  The surface emissivities are randomly chosen between 
0.8 and 1.0 and assumed constant for all radiometer frequencies.  Brightness temperatures are calculated at 
%89-, 183.31 ± 1.1, and 325.15 ± 1.5 GHz,
all frequencies of the 10 channels of the SAPHIR-NG 
radiometer envisioned to be deployed in the AOS mission \citep{brogniez2022time}.  One of the frequencies is 
89-GHz, while the others are centered on the 183.31, and 325.15 GHz water vapor absorption lines. Details regarding
the radiometer and the active instruments are
provided in Table \ref{t_sensors}.  Errors in the radiometer observations are modeled assuming a 
Noise-Equivalent-Delta-T (NEDT) of 1K, which is a readily achievable level for modern satellite radiometers \cite{draper2015}.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig06.rev.png}\\
    \caption{Same as in Figure 4, but for the IWC gravity center.}\label{f6}
\end{figure}

@@

Although retrievals from the lidar-only or radar-only observations are not expected to be as accurate as those from
radiometer-only observations, the lidar observations being subject to severe attenuation while the radar observations
being limited by its sensitivity, they are nevertheless insightful. This is because quantifying the limitations of
retrievals from active instruments may be used to better assess  the benefits of synergistic retrievals.  Shown in
top of row of Figure \ref{f7} are the distributions of IWP estimated from lidar-only and radar-only observations 
as a function of their true values. As apparent in the figure, and as expected, the lidar-only retrievals tend to
be accurate for IWP values smaller than 50 g/m$^2$, while the radar-only retrievals tend to be reliable only for large
IWP values on the order of hundreds of g/m$^2$.  Not that the radar-only retrievals exhibit a bimodal distributions,
with a broad distribution of real IWP values associated with a single retrieved IWP value.  This is a consequence
of the fact that there is a large number of atmospheric profiles characterized by not necessarily small IWP values that
are not associated with detectable radar signals. The results shown in the top row of Figure \ref{f7} are conditional on
either the CALIOP lidar or CS radar observations detecting ice, and so are the radar-only retrievals. However, the 
indiscriminate application of a Ku-band radar-only retrieval procedure to the entire dataset may result in significant
overestimation because clear skies would be associated with the same IWP value as ice clouds with reflectivity signal 
below the detection threshold. On the other hand, limiting the application of the radar-only retrieval procedure to atmospheric profiles
with detectable ice clouds would result in a significant underestimation, as a large number of atmospheric profiles
with ice clouds would be associated with zero IWP.  From this perspective, the derivation of  Ku-band radar-only ice estimates
is not useful if coincident radiometer observations are available.

>>>>

To achieve its objectives, AOS will rely on a combination of active and passive observations onboard multiple spacecrafts

@@

complimentary and enable the synergistic characterization of ice clouds. 

@@

good first step towards the development of an operational synergistic retrieval of ice clouds from AOS inclined observations.

@@

\subsection{Assumptions and forward models}
To quantify the number of ice particles in an elementary atmospheric volume as a function of their size, we use
normalized gamma functions \citep{bringi2003}.  The benefit of normalized gamma functions is that they 
encapsulate the variability of Ice Water Content (IWC) - reflectivity relationship into a single parameter,
i.e. the 
normalized Particle Size Distribution (PSD) intercept \citep{testud2001,bringi2003}. The normalized
PSD intercept is defined as $N_w=\frac {4^4} {\pi \rho_w} \frac {IWC} {D_m^4}$, where $IWC$ is the ice water 
content associated with the PSD, and $D_m$ is the mass weighted mean diameter.  Based on the work of \cite{testud2001},
\cite{ferreira2001} and \cite{delanoe2014} showed 
that the variability in IWC reflectivity (Z) relationships may be fully explained by variability in $N_w$, and 
that a formula of the type
\begin{equation}
IWC=N_w^{1-b}a_{norm} Z^b \label{eq:1}
\end{equation}
(where $a_{norm}$ and $b$ are constants) explains almost perfectly the relationships between IWC and Z calculated from observed PSDs.
However, equation (\ref{eq:1}) is not 
sufficient to derive accurate, unbiased estimates of ice water contents, because $N_w$ varies considerably in 
time and space. Nevertheless, multiple studies showed that it is beneficial to parameterize $N_w$ as a function 
of various variables, such as temperature \citep{hogan2006retrieval, delanoe2008variational,deng2010tropical}, 
rather than using $N_w$ independent reflectivity ice relations.  In this study, we parameterize
$N_w$ as a function of temperature based on the CloudSat 2C-ICE product \citep{deng2010tropical,deng2013evaluation}.  
A scatter-plot analysis of relationships between the 2C-ICE IWC and the associated reflectivity suggests that
the multiplicative coefficient in a power low ice-reflectivity relationship is parameterized as a function of temperature 
(or equivalently, as a function of height relative to the freezing level) in 
the default 2C-ICE retrievals. The default multiplicative coefficient (the value the provides the IWC estimate prior
to the ingestion of the lidar observations) may be simply estimated by regressing the 2C-ICE IWC on the associated CloudSat reflectivity 
as a function of height relative the freezing level. Specifically, a regression of the type $ln(IWC)=ln(a)+b*ln(Z)$ is performed for 
each bin referenced relative to the zero degree bin is performed, and the variation of $ln(a)$ is parameterized as a function of the
relative height.  The results are shown in Figure \ref{f2}. As apparent in the figure, and as expected, $a$ exhibits a strong variation 
with the relative height. Given that any deviation of the multiplicative coefficient in 
an IWC-Z relation from an average is equivalent to a deviation of the associated $N_w$ from its mean value 
\citep{ferreira2001,delanoe2014}, the variation of $a$ as a function of relative-height may be converted into an $N_w$ 
as a function of relative-height relationship.  We, therefore, use the data in Figure \ref{f2} to parameterize $N_w$ as a 
function of the relative height.  Specifically, given that $(1-b)ln(N_w)+ln(a_{norm})=ln(a)$ and $ln(a)=ln(a_0)+s*H$, with 
$a_0$ and $s$ the intercept and the slope of a regression for the data in Figure \ref{f2}, we 
can express $N_w$ as a function of the relative height, slope $s$, and parameters $a_0$ and $a_{norm}$.

@@

To investigate the variability in the vertical distribution of the 2C-ICE estimates and their consistency with our estimates, 
we cluster, based on similarity  (quantified through the evaluation of the Euclidian distance between profiles), 
a large set 2C-ICE profiles into 18 classes using a k-Means 
procedure. The mean IWC profiles associated with the 18 classes are shown in continuous lines in Figure \ref{f1}.  
Our estimates, derived using PSD assumptions and electromagnetic scattering calculations that enable 
accurate and physically consistent simulations of radar observations at Ku-band and radiometer observations of 
submillimeter-wave frequencies are also shown in the figure. These estimates are based on the self-similarity 
Rayleigh-Gans approximation (SSRGA) of \cite{hogan2017ssrga}. Details regarding the
estimation process are provided in the following paragraphs.  As apparent in Figure \ref{f1}, the CS and SSRGA estimates are in good 
agreement.  Some discrepancies due to differences between the SSRGA $N_w$ parameterization and the CS 2C-ICE 
"a priori assumptions" are also apparent, but they are not deemed critical in this study, whose objective is 
the investigation of synergistic lidar, Ku-band radar and submillimeter-wave radiometer retrievals, because the 
outcome is not likely to be sensitive to such details. Also apparent in Figure \ref{f1} is the fact that there is significant
variability in the vertical distribution of the 2C-ICE IWC estimates, which makes the estimation of IWC profiles from passive-only 
observations challenging.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig01.png}\\
    \caption{Mean CS IWC profiles for 18 classes derived using the k-Means clustering algorithm. For a 
    compressed but intelligible representation, three classes are shown in different colors in each panel.
    The associated 
    mean profiles derived from CS reflectivity observations using the SSRGA scattering calculations 
    and $N_w$ parameterization developed in this study are shown using symbol *. The vertical coordinate 
    is defined relative to the freezing level.}\label{f1}
\end{figure}

For the determination of reference $a_{norm}$ and $b$ values to be used with Equation (\ref{eq:1}), we assume that 
PSDs are normalized gamma  distributions with $N_w=0.08cm^{-4}$ and $\mu=2$ and calculate 
\begin{equation}
 Z=\frac {\lambda ^4} {\pi ^5 |K_w|^2} \int_0^{\infty} N(D,D_m) \sigma _b(D) dD
\end{equation}
where $\lambda$ is the radar frequency, $|K_w|$ is the dielectric factor of water, $N(D,D_m) dD$ is the number 
of ice particles of diameter with D and D+dD per unit volume, $D_m$ is the mass weighted mean diameter of the 
distribution, and $\sigma _b(D)$ is the backscattering cross-section of ice particle of diameter $D$. The mass 
weighted mean diameter is equidistantly sampled to span the entire range of IWC values in the CS 2C-ICE dataset.
The assumed mass-size relation is that of \cite{brown1995improved} because it works well with the 
SSRGA scattering calculations \citep{heymsfield2022impacts}.  The open source software scatter-1.1 of \citep{scatter-1.1}
is used to provide the actual scattering properties.   
To improve the representation of microphysical variability in the study, we do not assume the values of $N_w$ given
by the 2C-ICE based parametrization described above as the true values. Instead, we perturb them by multiplication 
with a log-normally distributed random variable with 0.0 mean a standard deviation of 0.5.  The 
perturbations are vertically correlated. Specifically, normal random variables with zero mean and a standard deviation
of 1.0 are generated and then passed through a Gaussian smoothing filter \citep{nixon2019} with a size of two radar bins. The perturbation
vector is then rescaled to have a standard deviation of 0.5.  The perturbations are then exponentiated and multiplied
with the 2C-ICE based $N_w$ values to produce the final $N_w$ values.  The filter size and noise magnitude are chosen
to roughly mimic the observed variability in $N_w$ derived from in situ observations of PSDs, such as those described
in \cite{heymsfield2022impacts}. In addition, we are adding a random noise of
0.0 mean and 0.5 dB standard deviation to the calculated reflectivity values, which is consistent with the expected
performance of a space-borne Ku-band radar system \citep{takahashi2008}.  Nevertheless, it should be noted that the
uncertainties in the reflectivity calculations are likely to be even larger, given that the SSRGA theory, although 
quite accurate, cannot possibly capture the entire range of uncertainties in the scattering properties of ice particles.
Specifically, the SSRGA calculations were carried out assuming ice particles consisting of aggregates of 
bullet rosettes, columnar crystals, and plates.

@@

The radiometer observations are calculated using a one-dimensional efficient, but accurate, radiative transfer 
solver based on Eddington's approximation \citep{kummerow1993edd}. The Eddington's approximation has been found to work 
well in cloud and precipitation retrieval application despite its simplicity relative to more general (but also 
computationally intensive) approaches such as Monte Carlo radiative transfer solvers \citep{liu1996three} 
It should be noted tough that the phase functions of ice particles tend to be highly asymmetric at 
sub-millimeter wave frequencies. For radiative transfer solutions based on the Eddington's approximation to be 
accurate, it is necessary that the delta-scaling approach \citep{joseph1976delta} be employed. The delta-scaling 
approach transforms the initial radiative transfer equation into an equivalent one characterized by a less 
asymmetric scattering function and more extinction, which makes the solution Eddington approximation more stable 
and accurate. The absorption due to water vapor and other gases is quantified using the Rosenkranz model 
\citep{rosenkranz1998}.  The water vapor, temperature, and pressure distributions are derived based on a WRF 
simulation of summer convection over the United States.  Specifically, the water vapor, temperature, and 
pressure profiles associated with times and areas where the model produces anvils are selected and clustered 
into 40 classes using the k-Means approach.  The mean extinction profiles at the radiometer frequencies are 
calculated for every class 
and used in process of calculating the brightness temperatures from the estimated ice profiles using a simple 
Monte Carlo procedure.  That is, given a retrieved ice profile and its scattering property, an anvil class 
and its associated absorption, temperature, and pressure profiles are randomly selected and attached to the 
ice scattering properties. To make the procedure physically meaningful, temperature rather than height is 
used in the ice scattering-gas absorption collocation process.  The surface emissivities are randomly chosen between 
0.8 and 1.0 and assumed constant for all radiometer frequencies.  Brightness temperatures are calculated at 
%89-, 183.31 ± 1.1, and 325.15 ± 1.5 GHz,
all frequencies of the 10 channels of the SAPHIR-NG 
radiometer envisioned to be deployed in the AOS mission \citep{brogniez2022time}.  One of the frequencies is 
89-GHz, while the others are centered on the 183.31, and 325.15 GHz water vapor absorption lines. Details regarding
the radiometer and the active instruments are
provided in Table \ref{t_sensors}.  Errors in the radiometer observations are modeled assuming a 
Noise-Equivalent-Delta-T (NEDT) of 1K, which is a readily achievable level for modern satellite radiometers \citep{draper2015}.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig06.rev.png}\\
    \caption{Same as in Figure \ref{f4}, but for the IWC gravity center.}\label{f6}
\end{figure}

@@

Although retrievals from the lidar-only or radar-only observations are not expected to be as accurate as those from
radiometer-only observations, the lidar observations being subject to severe attenuation while the radar observations
being limited by sensitivity, they are nevertheless insightful. This is because quantifying the limitations of
retrievals from active instruments may be used to better assess  the benefits of synergistic retrievals.  Shown in
top of row of Figure \ref{f7} are the distributions of IWP estimated from lidar-only and radar-only observations 
as a function of their true values. As apparent in the figure, and as expected, the lidar-only retrievals tend to
be accurate for IWP values smaller than 50 g/m$^2$, while the radar-only retrievals tend to be reliable only for large
IWP values on the order of hundreds of g/m$^2$.  Not that the radar-only retrievals exhibit a bimodal distributions,
with a broad distribution of real IWP values associated with a single retrieved IWP value.  This is a consequence
of the fact that there is a large number of atmospheric profiles characterized by not necessarily small IWP values that
are not associated with detectable radar signals. The results shown in the top row of Figure \ref{f7} are conditional on
either the CALIOP lidar or CS radar observations detecting ice, and so are the radar-only retrievals. However, the 
indiscriminate application of a Ku-band radar-only retrieval procedure to the entire dataset may result in significant
overestimation because clear skies would be associated with the same IWP value as ice clouds with reflectivity signal 
below the detection threshold. On the other hand, limiting the application of the radar-only retrieval procedure to atmospheric profiles
with detectable ice clouds would result in a significant underestimation, as a large number of atmospheric profiles
with ice clouds would be associated with zero IWP.  From this perspective, the derivation of  Ku-band radar-only ice estimates
is not useful if coincident radiometer observations are available.
